---
title: "Manuscript"
format: pdf
editor_options: 
  chunk_output_type: console
bibliography: resources/bib.bib
csl: resources/citation-style.csl
link-citations: true
---


```{r}
#| echo: false
#| warning: false
#| results: false
#| label: Required packages


library(tidyverse)
library(ggdag)
```



# 1. Introduction



# 2. Setting Up a Collaborative Workspace



# 3. Conducting Simulations Before Data Acquisition

Conducting simulations prior to data acquisition is useful for several reasons, including: 

1. *Model design*. By sampling from the prior distributions, we can understand our 
expectations about potential outcomes. This pre-data exploration offers insights 
into the implications of the prior assumptions.
2. *Model checking*. Once a model is updated using real data, simulating implied 
data can help assess the success of the fit and explore the model's behavior.
3. *Software validation*. In order to ensure that our model fitting software is
functioning correctly, it is helpful to simulate observations under a known model 
and then attempt to recover the parameter values from which the data were simulated. 
4. *Research design*. Simulating observations based on our hypothesis allows for 
an evaluation of the research design's effectiveness. This is similar to conducting 
a *power analysis*, but the possibilities are broader.   
5. *Forecasting*. Estimates derived from simulations can be utilized to generate 
predictions for new cases and future observations [@McElreath_2018].

To illustrate some of the properties of conducting simulations, we will consider 
the following hypothesis: **Increasing daily exercise time is associated with lower 
blood pressure levels in adults**. 


To investigate this hypothesis, it is valuable to invest effort into finding methods 
that distinguish causal inferences from associations. A helpful first step to do 
this is to construct a causal model that is separate from the statistical model. 
The simplest graphical representation of such a causal model is a *DIRECTED ACYCLIC GRAPH*, 
commonly referred to as a *DAG* [@McElreath_2018; @digitale_tutorial_2022].

A DAG primarily serves to represent our prior knowledge about biological and behavioral systems that may confound the specific causal research question. In our example, we aim to investigate the causal effect of increasing exercise time, denoted as "E", on blood pressure, "B". A basic DAG for this relationship might look like this (Figure 1):


```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "A DAG, assuming that the causal effect of increasing exercise time (E) on blood pressure (B) is not confounded."
#| label: Fig-dag1
#| fig-width: 6
#| fig-height: 3



coord <- tibble(name = c("E", "B"),
                x = c(1, 2),
                y = c(1, 1))

dag <- dagify(B ~ E,
              coords = coord)


fig1 <- dag |> 
  tidy_dagitty() |>
  ggplot(aes(x = x, y = y,
             xend = xend, yend = yend)) +
  geom_dag_point(color = "white") +
  geom_dag_text(color = "black", size = 8) +
  geom_dag_edges(arrow_directed = grid::arrow(length = grid::unit(5, "pt"), 
                                              type = "closed")) +
  
  theme_dag()



fig1

```


This DAG assumes that the causal effect of increasing exercise time on blood pressure 
is not confounded, which is arguably a naive assumption. For instance, both age and 
diet could potentially confound this relationship. A more comprehensive DAG that 
accounts for these confounders might appear as follows (Figure 2):


```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: "A DAG, assuming that the causal effect of increasing exercise time (E) on blood pressure (B) is confounded by both age (A) and diet (D)."
#| label: Fig-dag2
#| fig-width: 6
#| fig-height: 3



coord <- tibble(name = c("E", "B", "A", "D"),
                x = c(1, 2, 1.2, 1.7),
                y = c(1, 1, 2, 2))

dag <- dagify(B ~ E,
              B ~ A,
              E ~ A,
              B ~ D,
              coords = coord)


fig1 <- dag |> 
  tidy_dagitty() |>
  ggplot(aes(x = x, y = y,
             xend = xend, yend = yend)) +
  geom_dag_point(color = "white") +
  geom_dag_text(color = "black", size = 8) +
  geom_dag_edges(arrow_directed = grid::arrow(length = grid::unit(5, "pt"), 
                                              type = "closed")) +
  
  theme_dag()



fig1

```


In this adjusted DAG, we presume that age directly influences blood pressure— an 
assumption supported by the correlation between aging and increased blood pressures 
[@Miall_1967]. Additionally, age may indirectly influence exercise time since older 
individuals often engage in less physical activity. Diet is also included as a confounding 
factor affecting blood pressure, linked to dietary habits like high-fat intake [@Wilde_2000].

Given that we believe our DAG to be accurate, we can construct our model with stronger 
scientific justifications regarding which confounding factors should be accounted 
for and how we will do so.

**Model design** 

Having established the confounding factors of age and diet in the relationship between 
exercise time and blood pressure, we are now prepared to design our study. We plan 
to recruit 100 sedentary adults aged between 20 and 70 for a one-year training study, 
during which participants will adhere to the WHO guideline of 150 minutes of endurance 
exercise per week. Unfortunately, we will only be able to measure blood pressure 
after the intervention period due to logistical constraints. Therefore, we can only 
investigate whether there is an association between how many minutes of physical 
activity the participants complete and their blood pressure levels after the intervention. 
To control for age and diet, we will document participants' ages at the time of inclusion 
and assess their diets using a binomial scale, where 1 indicates a healthy diet and
0 indicates an unhealthy diet. 

The model can now be described mathematically, for example in terms of a simple 
linear regression model:


\begin{align}
\operatorname{BloodPressure}_i &= \beta_0 + \beta_1 \cdot \operatorname{Exercise}_i 
+ \beta_2 \cdot \operatorname{Age}_i + \beta_3 \cdot \operatorname{Diet}_i + \epsilon_i \\
\end{align}




Where: 

\begin{align}
\beta_0 &: \text{Intercept}\\
\beta_1 &: \text{Coefficient for exercise time}\\
\beta_2 &: \text{Coefficient for age}\\
\beta_3 &: \text{Coefficient for diet}\\
\epsilon_i &: \text{Random error term}\\
\end{align}





The linear regression model can be implemented in R as follows: 

```{r}
#| results: false
#| label: Example data and model

# Example dataset
data <- data.frame(
  exercise = rnorm(100, mean=150, sd=20),
  age = rnorm(100, mean=45, sd=15),
  diet = rnorm(100, mean=2, sd=1),
  bloodpressure = rnorm(100, mean=120, sd=10)
)

# Fit a simple linear regression model
model <- lm(bloodpressure ~ exercise + age + diet, data = data)

# Summarize the model results
summary(model)
```

**Run simulations**

We will now run simulations to evaluate our model design. First, we need to define 
our parameter values:


\begin{align}
\operatorname{BP}_i & = \beta_0 + \beta_E \cdot \operatorname{E}_i + \beta_A \cdot \operatorname{A}_i + \beta_D \cdot \operatorname{D}_i + \epsilon_i \\
\operatorname{E}_i & \sim \operatorname{Normal}(\mu_E, \sigma_E) \\
\operatorname{A}_i & \sim \operatorname{Normal}(\mu_A, \sigma_A) \\
\operatorname{D}_i & \sim \operatorname{Binomial}(n=1, p=0.5) \\
\epsilon_i & \sim \operatorname{Normal}(\mu_\epsilon, \sigma_\epsilon) \\
\beta_0 & = 120 \quad (\text{intercept}) \\
\beta_E & = -0.1 \quad (\text{beta for exercise}) \\
\beta_A & = 0.2 \quad (\text{beta for age}) \\
\beta_D & = -5 \quad (\text{beta for diet}) \\
\mu_E & = 150 \quad (\text{mean exercise}) \\
\mu_A & = 45 \quad (\text{mean age}) \\
\mu_\epsilon & = 0 \quad (\text{mean error term}) \\
\sigma_E & = 20 \quad (\text{exercise standard deviation}) \\
\sigma_A & = 10 \quad (\text{age standard deviation}) \\
\sigma_\epsilon & = 5 \quad (\text{error term standard deviation}) \\
\end{align}



After defining our parameter values, we can simulate data using these parameters. 
To do this, it is convenient to create a function where you specify information about 
each parameter that can be manipulated later, thereby enabling experimentation with 
different values to assess the model's robustness. Below is the code for this simulation:

```{r}
#| results: false
#| label: Sim.function


# Simulate data based on the hypothetical model

# A basic function
sim.fun <- function(n = 100,
                intercept = 120,
                mean_exercise = 150,
                beta_exercise = -0.1,
                sigma_exercise = 20,
                mean_age = 45,
                beta_age = 0.2,
                sigma_age = 10,
                size_diet = 1,
                prob_diet = 0.5,
                beta_diet = -5, 
                mean_epsilon = 0,
                sigma_epsilon = 5) {
  
  # This is the body 
  epsilon <- rnorm(n, mean = mean_epsilon, sd = sigma_epsilon) #error term 
  diet <- rbinom(n, size = size_diet, prob = prob_diet) #simulate diet 
  age <- rnorm(n, mean = mean_age, sd = sigma_age) #simulate age
  exercise <- rnorm(n, mean = mean_exercise, sd = sigma_exercise) #simulate exercise duration
  bloodpressure <- intercept + (beta_exercise * exercise) + (beta_age * age) + 
    (beta_diet * diet) + epsilon
  
  
  
  # The output
  return(data.frame(age, exercise, diet, bloodpressure, epsilon))
  
}


# Store the data frame as "sim_data"
sim_data <- sim.fun()


```

With our simulated dataset, we can now test our hypotheses using the linear model: 
*lm(bloodpressure ~ exercise + age + diet, sim_data)*. The output will indicate whether
there is an observed effect of exercise on blood pressure:

```{r}
#| echo: false
#| label: Simulatetd.m1

set.seed(1)
sim_data <- sim.fun()


sim.m1 <- lm(bloodpressure ~ exercise + age + diet, sim_data)

summary.m1 <- summary(sim.m1)

```



```{r}
#| echo: false
#| label: Sim.m1.output 


summary.m1

exercise_effect <- sprintf("%.3f",summary(sim.m1)$coefficients["exercise", "Estimate"])

```

In this model, we can see that each minute of increased exercise time is associated 
with a **`r exercise_effect`** mmHg lower blood pressure, when age and diet 
are included as covariates.

Now it is time to play with our simulation! Lets say we only manage to include 50 
participents, what will happen whit our results?

This can easily be tested by manipulating our function and then running the model, 
like this:


```{r}
#| label: Manipulate function 

# Changing number of observations to 50!
set.seed(7)
sim_data2 <- sim.fun(n = 50)


```



```{r}
#| echo: false
#| label: Sim.m2.output 


sim.m2 <- lm(bloodpressure ~ exercise + age + diet, sim_data2)

summary.sim.m2 <- summary(sim.m2)

summary.sim.m2

exercise_effect2 <- sprintf("%.3f", summary(sim.m2)$coefficients["exercise", "Estimate"])
```


The output for this model may reveal that the **`r exercise_effect2`** mmHg effect 
of exercise is no longer significant, suggesting that 50 participants may not be 
sufficient to detect the true effect of exercise which we know is **-0.1** mmHg. 
However, this is just one simulation; results may vary due to random chance. It is 
therefore prudent to conduct multiple simulations to evaluate the robustness of our 
model, which can be executed in a loop as demonstrated below:




```{r}
#| label: Simulations loop

set.seed(1)
results <- list()

for(i in 1:1000) {
  
  dat1 <- sim.fun()
  dat2 <- sim.fun(n = 50)
  
  
  m1 <- lm(bloodpressure ~ exercise + age + diet, data = dat1)
  m2 <- lm(bloodpressure ~ exercise + age + diet, data = dat2)

  
  results[[i]] <- data.frame(model = c("m1", "m2"),
             estimate = c(coef(m1)[2],
                          coef(m2)[2]))
  
  
  
}


```



To visualize the estimates of model 1 and model 2, it is convenient to create a 
figure similar to Figure 3.


```{r}
#| label: Simulation figure 
#| echo: false
#| fig-cap: "Demonstrating the variation in the estimated effect of exercise on blood pressure using 100 participants (m1) compared to 50 participants (m2) by simulating 1,000 different datasets."


bind_rows(results) |>
  ggplot(aes(model, estimate)) +
  geom_point(position = position_jitter(width = 0.1),
             alpha = 0.2, 
             shape = 21,
             color = "blue") + 
  theme_classic()
```


In the simulated models shown in Figure 3, the estimate for model m1 clusters more 
tightly around -0.1, which represents the true effect of exercise on blood pressure
as specified. In contrast, model m2 exhibits more variation, indicating that the 
effect is less reliably detected with only 50 participants.

This brief introduction illustrates how to conduct basic data simulations before 
data acquisition, empowering you to enhance the robustness of your research! 


# 4. Including Data Packages for Distribution



# 5. Creating Visualizations from Data Packages

# Bibliography